Latent Variable Clustering for Classification of Events in the AT-TPC

### 2020-05-25 doc/Manuscript/article.tex
Edits for Section 2

The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is very expensive computationally, because it involves the simulation of thousands of tracks for each recorded event, that need to be performed during the Monte-Carlo fitting sequence. Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computer cluster. In the case of an experiment where the reaction of interest would represent less than a few percents of the total cross section, this procedure would become highly inefficient and prohibitive. Most of the CPU time would be spent on simulating events that are not of interest. A ML algorithm able to classify the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts on events of interest only.

>>>>

The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence. Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster. In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive. Most of the CPU time would be spent on simulating events that are not of interest. An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest. In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
